---
title: "Meeting 8"
author: "PLS-PM Cross-validation and Simulated Data"
date: "04/27/2015"
output: beamer_presentation
---

```{r, echo=FALSE}
#Initialization

#Clear Environment
rm(list=ls())

#Load our Libraries
source("./lib/simplePLS.R")
source("./lib/graphUtils.R")

#Load Specific Code
source("./Meeting 8.R")

#Load Library for PCA
library(psych)

#Load Original Data
Anime=read.csv("./data/originalData.csv",header=T)

#Create Index to Randomize Data
set.seed(123)
index=sample.int(dim(Anime)[1],183,replace=F)

#Prepare variables for Cross-Validation Results
plsPredictTest <- NULL
plsResidualsTest <- NULL
lmPredictTest <- NULL
lmResidualsTest <- NULL
crossTrainData <-NULL
crossTestData <-NULL

#Run a cicle to take 10 subsamples of 10 observations each and run the models
for (i in seq(10,100, by =10)){
  
  #Get subsamples
  trainData <- Anime[-index[(i-9):i],]
  testData <- Anime[index[(i-9):i],]
  
  #Run The Models
  result<-runProcess(trainData,testData)

  #Get PLS Results
  plsPredictTest<-rbind(plsPredictTest,result$original$pls$predTest$predictedMeasurements)
  plsResidualsTest<-rbind(plsResidualsTest,result$original$pls$residualsTest)

  #Get Linear Model Results
  lmPredictTest<-rbind(lmPredictTest,result$original$lm$predTest$predictedMeasurements)
  lmResidualsTest<-rbind(lmResidualsTest,result$original$lm$residualsTest)
  
  #Save Actual Data
  crossTraintData <-rbind(crossTrainData,trainData)
  crossTestData <-rbind(crossTestData,testData)
}


```

## Cross-validation Method

This is the process performed to produce the Cross-validation Samples:

- Load the original 183 observations.
- Select a random sample of 183 observations from the original data (Randomize the Observations).
- The first 100 randomized observations were divided into 10 subsamples of 10 observations each one.
- Each subset of 10 observations was taken as a holdout for an individual model, and the remaining 173 observations as a training data.
- Each model produced its own predictions.
- All the predictions were combined together to produce a set of 100 predictions based on 10 different models, with different training and holdout data.


## Actual vs Predicted Scatterplot (PLS vs LM)

```{r, echo=FALSE}

#Set the panels
par(mfrow=c(2,2))

#PLS: Actual vs Predicted (AA.0)
y<-plsPredictTest[,"AA.0"]
x<-crossTestData[,"AA.0"]
z<-lmPredictTest[,"AA.0"]
title="Actual vs Predicted (AA.0)"
xlabel=paste("Mean: Act=",
             signif(mean(x),digits=4),
             "PLS=",
             signif(mean(y),digits=4),
             "LM=",
             signif(mean(z),digits=4),
             "\n SD: Act=",
             signif(sd(x),digits=4),
             "PLS=",
             signif(sd(y),digits=4),
             "LM=",
             signif(sd(z),digits=4))
ylabel="Predicted"
graphScatterplot(x,y,z,title,xlabel,ylabel)
points(x+0.1,z, col= "black")

#PLS: Actual vs Predicted (AA.1)
y<-plsPredictTest[,"AA.1"]
x<-crossTestData[,"AA.1"]
z<-lmPredictTest[,"AA.1"]
title="Actual vs Predicted (AA.1)"
xlabel=paste("Mean: Act=",
             signif(mean(x),digits=4),
             "PLS=",
             signif(mean(y),digits=4),
             "LM=",
             signif(mean(z),digits=4),
             "\n SD: Act=",
             signif(sd(x),digits=4),
             "PLS=",
             signif(sd(y),digits=4),
             "LM=",
             signif(sd(z),digits=4))
ylabel="Predicted"
graphScatterplot(x,y,z,title,xlabel,ylabel)
points(x+0.1,z, col= "black")

#PLS: Actual vs Predicted (AA.2)
y<-plsPredictTest[,"AA.2"]
x<-crossTestData[,"AA.2"]
z<-lmPredictTest[,"AA.2"]
title="Actual vs Predicted (AA.2)"
xlabel=paste("Mean: Act=",
             signif(mean(x),digits=4),
             "PLS=",
             signif(mean(y),digits=4),
             "LM=",
             signif(mean(z),digits=4),
             "\n SD: Act=",
             signif(sd(x),digits=4),
             "PLS=",
             signif(sd(y),digits=4),
             "LM=",
             signif(sd(z),digits=4))
ylabel="Predicted"
graphScatterplot(x,y,z,title,xlabel,ylabel)
points(x+0.1,z, col= "black")

#PLS: Actual vs Predicted (AA.3)
y<-plsPredictTest[,"AA.3"]
x<-crossTestData[,"AA.3"]
z<-lmPredictTest[,"AA.3"]
title="Actual vs Predicted (AA.3)"
xlabel=paste("Mean: Act=",
             signif(mean(x),digits=4),
             "PLS=",
             signif(mean(y),digits=4),
             "LM=",
             signif(mean(z),digits=4),
             "\n SD: Act=",
             signif(sd(x),digits=4),
             "PLS=",
             signif(sd(y),digits=4),
             "LM=",
             signif(sd(z),digits=4))
ylabel="Predicted"
graphScatterplot(x,y,z,title,xlabel,ylabel)
points(x+0.1,z, col= "black")
```

## PLS vs Linear Regression: Joint Histogram

```{r, echo=FALSE}

#Set the panels
par(mfrow=c(2,2))

title<-"PLS vs LM Residuals"
graphCombinedResiduals("AA.0",lmResidualsTest,plsResidualsTest,title,c(-6,6),c(0,0.4),10,"LM","PLS")

graphCombinedResiduals("AA.1",lmResidualsTest,plsResidualsTest,title,c(-6,6),c(0,0.4),10,"LM","PLS")

graphCombinedResiduals("AA.2",lmResidualsTest,plsResidualsTest,title,c(-6,6),c(0,0.4),10,"LM","PLS")

graphCombinedResiduals("AA.3",lmResidualsTest,plsResidualsTest,title,c(-6,6),c(0,0.4),10,"LM","PLS")
```
